---
title: "FundRmentals 06"
author: "Dr Danielle Evans | University of Sussex"
output:
  xaringan::moon_reader:
    css: ["xaringan-themer.css", "ppt_style.css"]
    lib_dir: libs
    nature:
      highlightStyle: monokai
      highlightLines: true
      highlightSpans: true
      ratio: '16:9'
---

```{r, echo=FALSE, eval = F}

xaringanExtra::use_logo("../img/logo.png", width = "30px", height = "30px", link_url = "https://fundrmentals.netlify.app/")

```

```{r setup, include=FALSE}
library(xaringanthemer)
library(xaringanExtra)
library(tidyverse)

options(htmltools.dir.version = FALSE)
xaringanExtra::use_clipboard()

```

```{r xaringan-themer, include=FALSE, warning=FALSE}

style_mono_light(
  base_color = "#23395b"
)

```


# Setup & Q&A

1. Open RStudio

1. Open/create your fundRmentals R Project (click the blue cube in the top right corner of RStudio)
  
1. Open the worksheet from last week, or [download it here...](https://fundrmentals.netlify.app/docs/week_04.Rmd) & save it to your **r_docs** folder

1. Run all the code chunks we created last week
<br>
<br>
<br>
<br>

.center[

### Any questions from the last tutorial?

]

---

# Overview

- Recap from last week

- The pipe **%>%** 

- Summary tables

- Next steps

---

# `dplyr::mutate()`

- `dplyr::mutate()` is for creating new columns/overwriting columns from existing ones by **mutating** them in some way

Creates a new column by multiplying the old values by 500: 

```{r, eval = F}
object_name <- dplyr::mutate(.data = object_name, new_column = old_column*500) 
```

Creates a new column by squaring the old values:

```{r, eval = F}
object_name <- dplyr::mutate(.data = object_name, new_column = old_column^2)  
```

Overwrites the old column by dividing the old values by 2:

```{r, eval = F}
object_name <- dplyr::mutate(.data = object_name, old_column = old_column/2)
```

<br>

**Task**: using the emp_data_2 object create a new column called **Attractiveness** from the **Q15620** column, by multiplying the scores by 100

---

# `dplyr::rename()`

- `dplyr::rename()` is for renaming columns

To rename one variable: 

```{r, eval = F}
object_name <- dplyr::rename(.data = object_name, new_column_name = old_column_name)
```

To rename multiple variables: 

```{r, eval = F}
object_name <- dplyr::rename(.data = object_name, 
                             new_column_name = old_column_name, 
                             new_column_name_2 = old_column_name_2, 
                             new_column_name_3 = old_column_name_3)
```

<br>
<br>
<br>
<br>

**Task**: using the emp_data_2 object rename the **Q14097** column to be called: **Empathy** 



---

# The Pipe %>% 

- The pipe **%>%** is part of the `magrittr` package & loads with `library(tidyverse)` or `library(magrittr)`

- We can use the pipe **%>%** to chain multiple commands together

- It pipes the output from the left into the functions/commands on the right: 

```{r, eval = F}
object_name %>% a_function() %>% another_function()
```

- It helps us avoid confusing nested functions which makes code easy to read & less prone to errors

- To insert a pipe **%>%** we can use the keyboard shortcut of **Ctrl + Shift + M** or **Command + Shift + M**

---

# The Pipe %>% 

**Nested example ()**

```{r, eval = FALSE}

am_routine <- leave_house(get_dressed(get_ready(wake_up(person = me, time = "too_early"),
  existential_crisis = TRUE, breakfast = FALSE, coffee_cups = 50), clothing = "pyjamas", 
  footwear = "fluffy_slippers"), university = FALSE, zoomiversity = TRUE)

```

<br>

**Piped example %>% **

```{r, eval = FALSE}

am_routine <- me %>%
    wake_up(person = ., time = "too_early") %>% 
    get_ready(person = ., existential_crisis = TRUE, breakfast = FALSE, coffee_cups = 50) %>% 
    get_dressed(person = ., clothing = "pyjamas", footwear = "fluffy_slippers") %>% 
    leave_house(person = ., university = FALSE, zoomiversity = TRUE)

```

- Whenever you see **.** or **.,** it's just a placeholder for the function argument that's being piped through


**Task**: try putting all the previous steps from today & last week (select, filter, mutate, rename) into one pipeline!

---

# `dplyr::summarise()`

- The summarise() function creates a summary table of our data

- Some common summary stats that we might use in it are mean(), median(), sd(), min(), max() & n()

- We specify the name we want to give our summary table object, the data, the column names for our summary table, & the summary stats we want to use with the column like this:

```{r, eval = FALSE}

object_name_summary <- object_name %>% 
  dplyr::summarise(
    mean_column = mean(column),
    sd_column = sd(column)
  )

```

- A common error is forgetting a comma between each summary stat or including one on the final line when we don't need to so watch out for this!

**Task**: add a new code chunk, & create a summary table of the mean, min & max values of Age from emp_data_3 - remember to type out the name of the summary table to see it!



???

im going to cover two functions we would often use with the pipe, the first being the summarise function from dplyr

---

# `dplyr::group_by()`

- The group_by() function allows us to group our summary further by some variable in our dataset

- You might want to see your summary stats split by different conditions, different age groups, different genders etc. (or all of these)

- To use it, we just add a line before the summarise() function: 

```{r, eval = FALSE}

object_name_summary <- object_name %>% 
  dplyr::group_by(some_grouping_variable) %>%
  dplyr::summarise(
    mean_column = mean(column),
    sd_column = sd(column)
  )

```

**Task**: edit the previous code chunk to split our Age summary table by Harm

**ExtRa Task**: try including two grouping variables by adding Humanness in the group_by() function along with Harm

---

# Next Steps

- Save the RMarkdown document from today **Ctrl + S** or **Command + S**

For next week you should complete the **discovr_07** tutorial on associations by pasting the below code into the **Console** in RStudio, pressing **Enter** 

```{r, eval = F, echo = T}

learnr::run_tutorial("discovr_07", package = "discovr")

```




