---
title: "FundRmentals 09"
author: "Dr Danielle Evans | University of Sussex"
output:
  xaringan::moon_reader:
    css: ["xaringan-themer.css", "ppt_style.css"]
    lib_dir: libs
    nature:
      highlightStyle: monokai
      highlightLines: true
      highlightSpans: true
      ratio: '16:9'
---

```{r, echo=FALSE, eval = F}

xaringanExtra::use_logo("../img/logo.png", width = "30px", height = "30px", link_url = "https://fundrmentals.netlify.app/")

```

```{r setup, include=FALSE}
library(xaringanthemer)
library(xaringanExtra)
library(tidyverse)

options(htmltools.dir.version = FALSE)
xaringanExtra::use_clipboard()
xaringanExtra::use_panelset()
```

```{r xaringan-themer, include=FALSE, warning=FALSE}

style_mono_light(
  base_color = "#23395b"
)

```


# Setup & Q&A

- Open RStudio

- Open your fundRmentals R Project (click the blue cube in the top right corner of RStudio)

- Open a new Rmd file for today

- In a new code chunk, load **tidyverse**, **palmerpenguins**, **broom**, **robust**, & **parameters**, using the **library()** command (install them in the console if you need to)

- In another code chunk, create **peng_data** by copying the below code:

```{r}
peng_data <- palmerpenguins::penguins
```

- Explore the data by using **View()** & **names()** on the **peng_data** object (in the console!!)

<br>

.center[

### Any questions from the last tutorial?

]

---

# Overview

- Linear model

  + 1 predictor: 'simple regression'

  + 2+ predictors: 'multiple regression'
  
  + Robust models

- Next steps

---

# lm() Function 

- We can use the **lm()** function to fit our model, the general structure is as follows:

```{r, eval = F, echo = T}
our_model <- lm(outcome ~ predictor(s), data = data, na.action = an action)
```

- Where we sub in the **name** we want to give **our_model**, the **outcome**, the **predictor(s)**, the **data**/**tibble**, & any actions we want to apply to missing data (i.e., **na.action = na.exclude**)

<br>

**Task**: Fit a linear model to our **peng_data**, using **body_mass_g** as the outcome, and **flipper_length_mm** as our predictor, give it the name **peng_model**

- We have some missing data today - make sure to exclude it by using **na.action = na.exclude**

<br>

--

**Solution**:

```{r}
peng_model <- lm(body_mass_g ~ flipper_length_mm, data = peng_data, na.action = na.exclude)
```

---

# Our Model

.pull-left[

To actually see the results, we can use **summary(peng_model)**

```{r, eval = F}
summary(peng_model)
```


.mini[

```{r, echo = F}
summary(peng_model)
```

]]

.pull-right[

Alternatively, we can use **broom::tidy()** & **broom::glance()** to look at our estimates & model fit 


```{r, eval = F}
broom::tidy(peng_model)
```

.mini[

```{r, echo = F}
broom::tidy(peng_model)
```

]

<br>

```{r, eval = FALSE}
broom::glance(peng_model)
```

.mini[

```{r, echo = F}
broom::glance(peng_model)
```

]]


---

# Multiple Predictors

- We can use the **lm()** function again to fit our model with multiple predictors, the general structure is as follows:

```{r, eval = F, echo = T}
our_model <- lm(outcome ~ predictor_1 + predictor_2, data = data, na.action = an action)
```

- Where we sub in the **name** we want to give **our_model**, the **outcome**, the **predictors** with a **+**, the **data**/**tibble**, & any actions we want to apply to missing data (i.e., **na.action = na.exclude**)

<br>

**Task 1**: Fit a linear model to our **peng_data**, using **body_mass_g** as the outcome, **flipper_length_mm** & **bill_length_mm** as the predictors, give it the name **peng_model_2**

- We have some missing data today - make sure to exclude it by using **na.action = na.exclude**

<br>

**Task 2**: view the results in the same way as before


---



**Model solution**:

.tiny[

```{r}
peng_model_2 <- lm(body_mass_g ~ flipper_length_mm + bill_length_mm, data = peng_data, na.action = na.exclude)
```

]

.pull-left[

**Results**:

```{r, eval = F}
summary(peng_model_2)
```


.mini[

```{r, echo = F}
summary(peng_model_2)
```

]]

.pull-right[

<br>

```{r, eval = F}
broom::tidy(peng_model_2)
```

.mini[

```{r, echo = F}
broom::tidy(peng_model_2)
```

]

<br>


```{r, eval = FALSE}
broom::glance(peng_model_2)
```

.mini[

```{r, echo = F}
broom::glance(peng_model_2)
```

]]

---

# Robust Models

For robust estimates we can use **robust::lmRob()** & **summary()** to see the output - like **lm()**: 

.tiny[

```{r, eval = F}
our_model_rob <- robust::lmRob(outcome ~ predictor_1 + predictor_2, data = data, na.action = an action)
summary(our_model_rob)
```

]

- We can then compare these to the original betas to see if our model was biased

<br>

For robust *p*-values & CIs we can use the **parameters::model_parameters()** function to estimate the model with standard errors designed for heteroscedastic residuals, where we give the name of our **original non-robust model**, set the **robust** argument to be **TRUE**, and the **vcov.type** to be **HC4**

```{r, eval = F}
parameters::model_parameters(non_robust_model, robust = TRUE, vcov.type = "HC4", digits = 3)
```


<br>

**Task**: fit our **peng_model_2** with robust estimates & robust CIs/*p*-values

---

# Solutions 

.mini[

```{r, warning=F, message=F}
peng_model_2_rob <- robust::lmRob(body_mass_g ~ flipper_length_mm + bill_length_mm, data = peng_data, na.action = na.exclude)

summary(peng_model_2_rob)
```

]

???

For robust parameters/estimates we can use the **robust::lmRob()** function & **summary()** to see the output (just like **lm()**): 

For robust p values & CIs we can use the **parameters::model_parameters()** function to estimate the model with standard errors designed for heteroscedastic residuals, here we give the name of our **original non-robust model**

---

# Solutions


```{r}
parameters::model_parameters(peng_model_2, robust = TRUE, vcov.type = "HC4", digits = 3)
```




???

For robust parameters/estimates we can use the **robust::lmRob()** function & **summary()** to see the output (just like **lm()**): 

For robust p values & CIs we can use the **parameters::model_parameters()** function to estimate the model with standard errors designed for heteroscedastic residuals, here we give the name of our **original non-robust model**



---

# Next Steps

Next week is 'reading week', feel free to use this to catch up on any tutorials you've missed so far

The next session in week 11, is 'pick a test week'! You can complete any of the below discovr tutorials, posting any questions or topic suggestions on [**this padlet**](https://uofsussex.padlet.org/de84/9gba8ktbw8egpqwm) & I'll cover anything you want in week 11 - do make sure to post your questions/comments as early as you can so that I can prepare anything I need to with enough time...

**discovr_10** - moderation & mediation  
**discovr_11** - ANOVA  
**discovr_12** - ANCOVA   
**discovr_13** - factorial designs   
**discovr_14** - repeated measures designs  
**discovr_15** - mixed designs  
**discovr_17** - exploratory factor analysis  
**discovr_20** - logistic regression  

Paste the below code into the **Console** in RStudio, change it to the **tutorial name** you want to do, and press **Enter** 

```{r, eval = F, echo = T}

learnr::run_tutorial("tutorial_name", package = "discovr")

```






