---
title: "FundRmentals 04"
author: "Dr Danielle Evans | University of Sussex"
output:
  xaringan::moon_reader:
    css: ["xaringan-themer.css", "ppt_style.css"]
    lib_dir: libs
    nature:
      highlightStyle: monokai
      highlightLines: true
      highlightSpans: true
      ratio: '16:9'
---

```{r, echo=FALSE, eval = F}

xaringanExtra::use_logo("../img/logo.png", width = "30px", height = "30px", link_url = "https://fundrmentals.netlify.app/")

```

```{r setup, include=FALSE}
library(xaringanthemer)
library(xaringanExtra)
library(tidyverse)

options(htmltools.dir.version = FALSE)
xaringanExtra::use_clipboard()

```

```{r xaringan-themer, include=FALSE, warning=FALSE}

style_mono_light(
  base_color = "#23395b"
)

```


# Setup & Q&A

1. Open RStudio

1. Open/create your fundRmentals R Project (click the blue cube in the top right corner)

  + To create one, go to **File**, **New Project**, **New Directory**, **New Project**, give your project a name, then click **Create**
  + Then create **data** & **r_docs** folders (in the **Files pane** click **New Folder**)
  
1. [Download the worksheet here...](https://fundrmentals.netlify.app/docs/week_04.Rmd) & save it to your **r_docs** folder

<br>
<br>
<br>
<br>
<br>

.center[

### Any questions from the last tutorial?

]

---

# Overview

- `tidyverse`

- Wrangling data

- The pipe **%>%** 

- Next steps

---

# `tidyverse` & `dplyr`

- The `tidyverse` is a collection of packages that all share a similar design philosophy, grammar, & data structure

- When we load `tidyverse`, it loads several different packages automatically:

  + `readr`, `dplyr`, `forcats`, `ggplot2`, `tidyr`, `purrr`, `tibble`, & `stringr`

- `dplyr` (like pliers..) is a useful package for wrangling/manipulating data & includes functions to:
  
  + `select()`,
  
  + `filter()`,
  
  + `mutate()` & 
  
  + `rename()` variables 

---

# `dplyr::select()`

- `dplyr::select()` is for selecting or removing specific columns in a dataset, follows this structure:

```{r, eval = F}
object_name <- dplyr::select(.data = object_name, column_name) # for selecting
```

```{r, eval = F}
object_name <- dplyr::select(.data = object_name, -column_name) # for removing
```

#### Assignment <- 

- Last week we learned that we can save the output of commands by creating an object with the assignment operator **<-**

- When manipulating our data sometimes it might be useful to create a new object (& keep the original) by using a different object name, and other times it might be more useful to overwrite the original object by using the same object name

**Task**: remove the **Experience** column & create a new object by calling it **emp_data_2**


---

# `dplyr::filter()`

- `dplyr::filter()` is for **keeping** rows of data based on some conditions like these:

.tiny[

To keep rows equal to some text:

```{r, eval = F}
object_name <- dplyr::filter(.data = object_name, column_name == "some text") 
```

To keep rows not equal to some text: 

```{r, eval = F}
object_name <- dplyr::filter(.data = object_name, column_name != "some text") 
```

To keep rows where value is smaller than 50:

```{r, eval = F}
object_name <- dplyr::filter(.data = object_name, column_name < 50)
```

To keep rows that meet BOTH conditions;

```{r, eval = F}
object_name <- dplyr::filter(.data = object_name, column_name_1 < 2 & column_name_2 == FALSE) 
```

]

**Task**: filter the data by keeping participants above **Age 29**

---

# `dplyr::mutate()`

- `dplyr::mutate()` is for creating new columns/overwriting columns from existing ones by **mutating** them in some way

Creates a new column by multiplying the old values by 500: 

```{r, eval = F}
object_name <- dplyr::mutate(.data = object_name, new_column = old_column*500) 
```

Creates a new column by squaring the old values:

```{r, eval = F}
object_name <- dplyr::mutate(.data = object_name, new_column = old_column^2)  
```

Overwrites the old column by dividing the old values by 2:

```{r, eval = F}
object_name <- dplyr::mutate(.data = object_name, old_column = old_column/2)
```

<br>

**Task**: create a new column called **Attractiveness** from the **Q15620** column, by multiplying the scores by 100

---

# `dplyr::rename()`

- `dplyr::rename()` is for renaming columns

To rename one variable: 

```{r, eval = F}
object_name <- dplyr::mutate(.data = object_name, new_column_name = old_column_name)
```

To rename multiple variables: 

```{r, eval = F}
object_name <- dplyr::mutate(.data = object_name, 
                             new_column_name = old_column_name, 
                             new_column_name_2 = old_column_name_2, 
                             new_column_name_3 = old_column_name_3)
```

<br>
<br>
<br>
<br>

**Task**: rename the **Q14097** column to be called: **Empathy** 



---

# The Pipe %>% 

- The pipe **%>%** is part of the `magrittr` package & loads with `library(tidyverse)` or `library(magrittr)`

- We can use the pipe **%>%** to chain multiple commands together

- It pipes the output from the left into the functions/commands on the right: 

```{r, eval = F}
a_function(object_name) %>% another_function()
```

- It helps us avoid confusing nested functions which makes code easy to read & less prone to errors

- To insert a pipe **%>%** we can use the keyboard shortcut of **Ctrl + Shift + M** or **Command + Shift + M**

---

# The Pipe %>% 

**Nested example ()**

```{r, eval = FALSE}

am_routine <- leave_house(get_dressed(get_ready(wake_up(person = me, time = "too_early"),
  existential_crisis = TRUE, breakfast = FALSE, coffee_cups = 50), clothing = "pyjamas", 
  footwear = "fluffy_slippers"), university = FALSE, zoomiversity = TRUE)

```

<br>

**Piped example %>% **

```{r, eval = FALSE}

am_routine <- me %>%
    wake_up(person = ., time = "too_early") %>% 
    get_ready(person = ., existential_crisis = TRUE, breakfast = FALSE, coffee_cups = 50) %>% 
    get_dressed(person = ., clothing = "pyjamas", footwear = "fluffy_slippers") %>% 
    leave_house(person = ., university = FALSE, zoomiversity = TRUE)

```

- Whenever you see **.** or **.,** it's just a placeholder for the function argument that's being piped through


**Task**: try putting all the previous steps (select, filter, mutate, rename) into one pipeline!

---

# Next Steps

- Save the RMarkdown document from today **Ctrl + S** or **Command + S**

- For next week you should complete the **discovr_05** tutorial by pasting the below code into the **Console** in RStudio, pressing **Enter** 

```{r, eval = F, echo = T}

learnr::run_tutorial("discovr_05", package = "discovr")

```

- This tutorial is pretty long, and generally this isn't something we learn by heart (or need to), you should aim to try go through the ggplot2 section as a minimum, then ideally complete the boxplots, plotting means, and scatterplots sections (you can ignore the faceting tasks if you want to)
  + There's also an empty code chunk for you to have a go at creating a plot from the emp_data in the worksheet for today

- Next week we'll cover summarising/visualising data with some RMarkdown thrown in for fun!






